---
format: 
  html: default
  revealjs: 
    output-file: 2_2_quarto_presentation.html
---

# Quarto

## What is Quarto?

"**An open-source scientific and technical publishing system"**

Works seamlessly within RStudio

Next generation version of R Markdown.

### Formats

-   Single documents

    -   PDF article

    -   HTML article

    -   HTML presentation

-   Collection of documents

    -   Book (html and/or PDF)

    -   Website (html pages, navigation)

## Benefits of using Quarto over just R.

## Working with Quarto documents

### Typing text

Source vs Visual

### Code chunks

```{r}

```

### Rendering

All code gets executed from scratch

## Getting started

### A single document

File \> New File \> Quarto Document

YAML section

``` yaml
---
title: "Untitled"
---
```

Text.

Code chunks.

```{r}
# hello!
```

### A Quarto Project

#### Project Features

The `_quarto.yml` file.

Contains rendering instructions for the entire project, and things that will apply to each file.

``` {.yaml filename="_quarto.yml"}
project:
  type: website

format:
  html:
    toc: true
```
